---
title: "Birmingham Pharmacy Access"
author: "David Ellis"
format: 
  html:
    embed-resources: true
    css: styles.css
    #bcc-html: default
editor: source
echo: false
warning: false
---

```{r setup}
library(readxl)
library(dplyr)
library(janitor)
library(leaflet)
library(htmltools)
library(sf)
library(BSol.mapR)
library(wmSpatial)
library(plotly)
source("functions.R")
source("api-functions.R")
source("configs.R")
# 1 mile in km
key_distance_km = 1.609344

# Variable access radius array
radii_list <- 1.609344 * c(0.25, 0.5, 1, 2) 
```


## Current Pharmacy Provision

```{r load-pharm-data}

pharm_data <- get_pharmacy_data("Most recent") %>%
  clean_names() %>%
  rename(
    postcode = post_code,
    name = organisation_name) %>%
  select(
    postcode, name, health_and_wellbeing_board, contains("opening_hours"), contract_type, period
  ) %>%
  mutate(
    popup_text = paste0(
      "<b>", name, "</b><br><br>",
      "<b>Monday</b>:", pharmacy_opening_hours_monday, "<br>",
      "<b>Tuesday</b>:", pharmacy_opening_hours_tuesday, "<br>",
      "<b>Wednesday</b>:", pharmacy_opening_hours_wednesday, "<br>",
      "<b>Thursday</b>:", pharmacy_opening_hours_thursday, "<br>",
      "<b>Friday</b>:", pharmacy_opening_hours_friday, "<br>",
      "<b>Saturday</b>:", pharmacy_opening_hours_saturday, "<br>",
      "<b>Sunday</b>:", pharmacy_opening_hours_sunday
    ),
    health_and_wellbeing_board = stringr::str_to_title(health_and_wellbeing_board)
  )%>%
  filter(
    # Filter out distance prescribing pharmacies
    contract_type != "DAC",
    postcode %in% wm_postcodes$postcode
  )



eth_data <- read_excel(GP_data_path)

latest_date <- pharm_data$period[[1]]
```
:::{.panel-tabset}

## All Days

```{r map-pharm-locs-all-days}
alldays_access <- basic_access_analysis(
    pharm_data,
    radius_km = key_distance_km,
    day = "All Days",
    radii_km_list = radii_list,
    eth_data = eth_data
  )

```
In `r latest_date` there were `r alldays_access$open_count` pharmacies in Birmingham. An estimated `r alldays_access$access_perc`% of Birmingham residents live within 1 mile (14 - 20 minutes walking distance) of a pharmacy that is open at least one day of the week. It should be noted that pharmacy access near the outer Birmingham border may be underestimated since pharmacies outside Birmingham are not included.

```{r}
#| fig-width: 6
#| fig-height: 5.5

alldays_access$m
```

An estimated `r round(alldays_access$var_dist$access_perc[1], 1)`% and `r round(alldays_access$var_dist$access_perc[2], 1)`% of Birmingham residents live within `r round(alldays_access$var_dist$radii_miles[1], 1)` miles and `r round(alldays_access$var_dist$radii_miles[2], 1)` miles respectively of a pharmacy that is open at least one day of the week.

```{r}
#| fig-width: 6
#| fig-height: 3

ggplotly(alldays_access$var_dist_plot)
```

Overall pharmacy access varies very little across ethnicity. The estimated proportion of residents from each of the 2021 census ethnic groups living within 1 mile of a pharmacy varies from `r min(alldays_access$eth_access$access_perc)`% to `r max(alldays_access$eth_access$access_perc)`%.

```{r}
#| fig-width: 6
#| fig-height: 4

ggplotly(alldays_access$eth_plot)
```


## Saturday

```{r map-pharm-locs-sat}
saturday_access <- basic_access_analysis(
    pharm_data,
    radius_km = key_distance_km,
    day = "Saturday",
    radii_km_list = radii_list,
    eth_data = eth_data
  )

sat_highest <- saturday_access$eth_access %>%
    slice_max(access_perc, n = 2, with_ties = FALSE)

sat_lowest <- saturday_access$eth_access %>%
    slice_min(access_perc, n = 2, with_ties = FALSE)

```

There are currently `r alldays_access$open_count` pharmacies in Birmingham. Of these, `r saturday_access$open_count` (`r round(100*saturday_access$open_count/nrow(pharm_data),1)`%) are open on Saturday. An estimated `r saturday_access$access_perc`% of Birmingham residents live within 1 mile (14 - 20 minutes walking distance) of a pharmacy that is open on Saturdays. It should be noted that pharmacy access near the outer Birmingham border may be underestimated since pharmacies outside Birmingham are not included.

```{r}
#| fig-width: 6
#| fig-height: 5.5

saturday_access$m
```

An estimated `r round(saturday_access$var_dist$access_perc[1], 1)`% and `r round(saturday_access$var_dist$access_perc[2], 1)`% of Birmingham residents live within `r round(saturday_access$var_dist$radii_miles[1], 1)` miles and `r round(saturday_access$var_dist$radii_miles[2], 1)` miles respectively of a pharmacy that is open on Saturdays.

```{r}
#| fig-width: 6
#| fig-height: 3

ggplotly(saturday_access$var_dist_plot)
```

Estimated pharmacy access on a Saturday shows limited variation across ethnicity. The ethnic groups with the estimated proportion living within 1 mile of a pharmacy open on a Saturday are `r sat_highest$ethnicity[1]` and `r sat_highest$ethnicity[2]` with `r sat_highest$access_perc[1]`% and `r sat_highest$access_perc[2]`% respectively. The groups with the lowest estimated access are `r sat_lowest$ethnicity[1]` and `r sat_lowest$ethnicity[2]` with `r sat_lowest$access_perc[1]`% and `r sat_lowest$access_perc[2]`% respectively.

```{r}
#| fig-width: 6
#| fig-height: 4

ggplotly(saturday_access$eth_plot)
```

## Sunday

```{r map-pharm-locs-sun}

sunday_access <- basic_access_analysis(
    pharm_data,
    radius_km = key_distance_km,
    day = "Sunday",
    radii_km_list = radii_list,
    eth_data = eth_data
  )

sun_highest <- sunday_access$eth_access %>%
    slice_max(access_perc, n = 2, with_ties = FALSE)

sun_lowest <- sunday_access$eth_access %>%
    slice_min(access_perc, n = 2, with_ties = FALSE)

```

There are currently `r alldays_access$open_count` pharmacies in Birmingham. Of these, `r sunday_access$open_count` (`r round(100*sunday_access$open_count/nrow(pharm_data),1)`%) are open on Sunday. An estimated `r sunday_access$access_perc`% of Birmingham residents live within 1 mile (14 - 20 minutes walking distance) of a pharmacy that is open on Sundays. It should be noted that pharmacy access near the outer Birmingham border may be underestimated since pharmacies outside Birmingham are not included.

```{r}
#| fig-width: 6
#| fig-height: 5.5

sunday_access$map
```


An estimated `r round(sunday_access$var_dist$access_perc[1], 1)`% and `r round(sunday_access$var_dist$access_perc[2], 1)`% of Birmingham residents live within `r round(sunday_access$var_dist$radii_miles[1], 1)` miles and `r round(sunday_access$var_dist$radii_miles[2], 1)` miles respectively of a pharmacy that is open on Sundays.

```{r}
#| fig-width: 6
#| fig-height: 3

ggplotly(sunday_access$var_dist_plot)
```

Estimated pharmacy access on a Sunday shows moderate variation across ethnicity. The ethnic groups with the estimated proportion living within 1 mile of a pharmacy open on a Sunday are `r sun_highest$ethnicity[1]` and `r sun_highest$ethnicity[2]` with `r sun_highest$access_perc[1]`% and `r sun_highest$access_perc[2]`% respectively. The groups with the lowest estimated access are `r sun_lowest$ethnicity[1]` and `r sun_lowest$ethnicity[2]` with `r sun_lowest$access_perc[1]`% and `r sun_lowest$access_perc[2]`% respectively.


```{r}
#| fig-width: 6
#| fig-height: 4

ggplotly(sunday_access$eth_plot)
```

:::

## Change Since Last Year

```{r load-last-year}

last_year_data <- get_pharmacy_data("Last year") %>%
  clean_names() %>%
  rename(
    postcode = post_code,
    name = organisation_name) %>%
  select(
    postcode, name, health_and_wellbeing_board, contains("opening_hours"), contract_type, period
  ) %>%
  mutate(
    popup_text = paste0(
      "<b>", name, "</b><br><br>",
      "<b>Monday</b>:", pharmacy_opening_hours_monday, "<br>",
      "<b>Tuesday</b>:", pharmacy_opening_hours_tuesday, "<br>",
      "<b>Wednesday</b>:", pharmacy_opening_hours_wednesday, "<br>",
      "<b>Thursday</b>:", pharmacy_opening_hours_thursday, "<br>",
      "<b>Friday</b>:", pharmacy_opening_hours_friday, "<br>",
      "<b>Saturday</b>:", pharmacy_opening_hours_saturday, "<br>",
      "<b>Sunday</b>:", pharmacy_opening_hours_sunday
    ),
    health_and_wellbeing_board = stringr::str_to_title(health_and_wellbeing_board)
  )%>%
  filter(
    # Filter out distance prescribing pharmacies
    contract_type != "DAC",
    postcode %in% wm_postcodes$postcode
  )

last_year_date <- last_year_data$period[[1]]

lastyear_allday_access <- basic_access_analysis(
    last_year_data,
    radius_km = key_distance_km,
    day = "All Days",
    radii_km_list = radii_list,
    eth_data = eth_data
  )

access_change <- alldays_access$var_dist$access_perc[[3]] - lastyear_allday_access$var_dist$access_perc[[3]]
  
increase_decrease <- ifelse(
  access_change > 0, "increase", "decrease"
)
```

In `r last_year_date` there were `r lastyear_allday_access$open_count` pharmacies in Birmingham. An estimated `r lastyear_allday_access$access_perc`% of Birmingham residents live within 1 mile (14 - 20 minutes walking distance) of a pharmacy that is open at least one day of the week. It should be noted that pharmacy access near the outer Birmingham border may be underestimated since pharmacies outside Birmingham are not included. Since last year, there has been a `r round(abs(access_change),2)`% `r increase_decrease` in proportion of residents living within one mile of a pharmacy.